Discrete Ordinates

### 2017-04-24 docprep/chapmeth.tex
Updated the thesis chapters and the beamshapes figures

The $S_N$ quadrature is implemented in DOCTORS. There are many different quadrature sets, the $S_N$ quadratures were selected because they are easy to implement, rotationally symmetric, and the most commonly used in production discrete ordinate solvers. This enables simple, direct comparison to other solvers. Arbitrary quadratures are permissible in a plain text file supplied by the user enabling other quadratures.

@@

@@

Voxel indices are flattened according to the rule given in Eq.~\ref{eq:indx_flat}. Note that all indeces always start at zero.

@@

\begin{equation} \label{eq:kronecker}
\delta(x, y) = 
\begin{cases}
1, \quad \mathrm{if} \ x=y \\
0, \quad \mathrm{otherwise}
\end{cases}
\end{equation}

@@

Substituting Eq.~\ref{eq:uncsrc} into $S_u$ in Eq.~{eq:col} gives Eq.~\ref{eq:col2} which simplifies to Eq.~\ref{eq:col3} by using Eq.~\ref{eq:totflux} to replace the collided and uncollided components with the total angular flux.

@@

>>>>

The rotation symmetrical quadrature is implemented in DOCTORS. There are many different quadrature sets, the rotation symmetrical quadratures were selected because they are easy to implement, rotationally symmetric, and the most commonly used in production discrete ordinate solvers. This enables simple, direct comparison to other solvers. Arbitrary quadratures are permissible in a plain text file supplied by the user enabling other quadratures.

@@

\subsection{Cross Section Parsing}
AMPX cross sections are read in binary. They are formatted like so. They are read like so. They were obtained from SCALE6.2. These files can be generated by other codes.

@@

Voxel indices are reordered into a one dimensional vector (flattened) according to the rule given in Eq.~\ref{eq:indx_flat}. Note that indices start at zero.

@@

\begin{equation} \label{eq:kronecker}
\delta(\vec{A}, \vec{B}) = 
\begin{cases}
1, \quad \mathrm{if} \ \vec{A}=\vec{B} \\
0, \quad \mathrm{otherwise}
\end{cases}
\end{equation}

@@

Substituting Eq.~\ref{eq:uncsrc} into $S_u$ in Eq.~\ref{eq:col} gives Eq.~\ref{eq:col2} which simplifies to Eq.~\ref{eq:col3} by using Eq.~\ref{eq:totflux} to replace the collided and uncollided components with the total angular flux.

@@

\section{Source Generation}
There are many kinds of external sources implemented in this work.

\subsection{Point Source}
The simplest source implemented is an isotropic point source. This source is defined only by its position and energy distribution. The center of the cell whose flux is being computed is located at position $<C_x, C_y, C_z>$ and the point source is located at position $<S_x, S_y, S_z>$.

\subsection{Fan Source}
In addition to the values defining a point source (position $\vec{S}$ and the energy distribution) a fan source is defined by two additional parameters, $\varphi$ and $\theta$ which describe the azimuthal and polar angles subtented by the fan beam respectively. Figure~\ref{fig:fanbeam} shows $\varphi$ and $\theta$ in reference to the geometry setup. The raytracing algorithm used for the fan beam is identical to the point source raytracer except for one difference: before the raytracing is done, whether the ray falls within the fan beam is determined and any ray falling outside the fan beam is not traced, its uncollided flux is immediately set to zero.

Some voxels would be partially enclosed inside the beam. A partial acceptance value is computed for these voxels and multiplied by the final result of the raytrace.

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/fan_rejection}
  \end{center}
  \caption{The fan.}
\label{fig:fan_rejection}
\end{figure}

First, the ray direction is projected onto the $xy$ plane giving $<S_x, S_y, 0>$ and $C_x, C_y, 0$. Then the vector from the source to the geometry center and the vector from the source to the cell center are dotted to give the cosine of the angle between them as shown in Eq.~\ref{eq:phicos}. Once $\zeta$ is computed, the condition given by Eq.~\ref{eq:phicoscon} is evaluated. If the condition is met, the angle is accepted and raytraced normally, otherwise, it is rejected immediately. Figure~\ref{fig:phi}

\begin{figure}[tb]
  \begin{center}
   \includegraphics[width=3.75in]{figs/phi}
  \end{center}
  \caption{How $\varphi$ is computed.}
\label{fig:phi}
\end{figure}

\begin{equation}\label{eq:phicos}
\cos(\zeta_\varphi) = \frac{<S_x, S_y, 0>}{\sqrt{S_x^2 + S_y^2}} \cdot \frac{<C_x, C_y, 0>}{\sqrt{C_x^2 + C_y^2}} = \frac{S_x C_x + S_y C_y}{\sqrt{S_x^2 + S_y^2} \sqrt{C_x^2 + C_y^2}}
\end{equation}

\begin{equation}\label{eq:phicoscon}
|\zeta_\varphi| < \frac{\varphi}{2}
\end{equation}

Treatment for the $\theta$ variable is similar to that of $\varphi$ except the polar angle is required instead of the azimuthal angle. This is computed using Eq.~\ref{eq:thetacos} for an arbitrary vector $\vec{\xi}$. This is used to compute the polar angle for both $\vec{S}$ and $\vec{C}$. The difference between the polar angles of the two vectors is computed using Eq.~\ref{eq:thetacos2} and then the condition given in Eq.~\ref{eq:thetacoscon} is evaluated. If the condition is not met, the ray is rejected and no raytrace is performed.

\begin{equation}\label{eq:thetacos}
\cos(\theta_\xi) = \frac{\xi_z}{\sqrt{\xi_z^2 + \xi_y^2 + \xi_z^2}}
\end{equation}

\begin{equation}\label{eq:thetacos2}
\zeta_\theta = \theta_S - \theta_C
\end{equation}

\begin{equation}\label{eq:thetacoscon}
|\zeta_\theta| < \frac{\vartheta}{2}
\end{equation}

\subsection{Multi-fan Source}
A focal point is determined. The focal spot is at the center of each fan beam. The focal spot is always assumed to be at the center of the geometry and at the same $z$ elevation as the source point.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/proj6}
        \caption{$N=6$}
        \label{fig:proj6}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/proj12}
        \caption{$N=12$}
        \label{fig:proj12}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
    %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/proj16}
        \caption{$N=16$}
        \label{fig:proj16}
    \end{subfigure}
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{figs/proj24}
        \caption{$N=24$}
        \label{fig:proj24}
    \end{subfigure}
    \caption{Different numbers of fan beam projections.}\label{fig:fanproj}
\end{figure}

\begin{equation}\label{eq:xp}
x' = x \cos \theta - y \sin \theta
\end{equation}

\begin{equation}\label{eq:yp}
y' = y \cos \theta + x \sin \theta
\end{equation}

To rotate a point $(x,y)$ about an arbitrary point $(x_f, y_f)$, Eq.~\ref{eq:xp2} and~\ref{eq:yp2} are used.

\begin{equation}\label{eq:xp2}
x' = (x-x_f) \cos \theta - (y-y_f) \sin \theta + x_f
\end{equation}

\begin{equation}\label{eq:yp2}
x' = (y - y_f) \cos \theta + (x - x_f) \sin \theta + y_f
\end{equation}

\subsection{Cone Source}
Pointy!

\subsection{Multi-cone Source}
All of the pointy!

\subsection{Multileaf Collimator}
How fancy.
